Design Document: Monte Carlo Maximum Likelihood for Generalized Linear Mixed Models

### 2014-11-24 NewDesignDoc/DesignDocGLMM.tex
Completed update to CLT proof

\section{Summary of model}
Typing \texttt{summary(mod)}  gives more detailed info of the model. This is broken into two pieces (as is usually done for summaries): there will be the \texttt{summary.glmm} and \texttt{print.summary.glmm}. We split this up because we have a lot of information in the summary that we do not necessarily want printed each time. When a user types \texttt{summary(mod)}, only the  basic information is automatically printed. More information can be found in the summary list.

The summary will do all the calculations and its value will be a list of the following:
\begin{itemize}
\item call
\item the variance estimate(s)
\item a matrix with the predictor in the first column, the estimated coefficient in the second column, the standard error in the third column, the z value in the fourth column, and the two-sided pvalue  in the fifth column. All inference is asymptotic, so we use the standard normal distribution to calculate the p-values.
\item the evaluated Monte Carlo log likelihood along with its first and second derivative
\item maybe other things that I haven't thought of
\end{itemize}

@@

Third and lastly, we want to show that 
\begin{align}
\sum_{k=1}^m \nabla \log \tilde{f} (u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}
\end{align}
 has finite variance. That is, we want to show the existence of
\begin{align}
\int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du.
\end{align}
We see that
\begin{align}
\int& \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du\\
&\leq  \int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)} du\\
&= \dfrac{1}{p_1} \int N(u|0,D) \left[\nabla \log \tilde{f}u) \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)+p_2 N(u|u^*,D^*)+p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})}   \right]^2 du\\
&\leq\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)}   \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{ \dfrac{\partial}{\partial \nu} N(u|0,D)}{ N(u|0,D)}   \right]^2 du\\
&= \dfrac{1}{p_1} \int \dfrac{\left[ \dfrac{\partial}{\partial \nu} N(u|0,D)  \right]^2}{N(u|0,D)} du \\
\end{align}

>>>>

\section{Summary of model}
Typing \texttt{summary(mod)}  gives more detailed info of the model. This is broken into two pieces (as is usually done for summaries):  \texttt{summary.glmm} and \texttt{print.summary.glmm}. We split this up because we have a lot of information in the summary that we do not necessarily want printed each time. When a user types \texttt{summary(mod)}, only the  basic information is automatically printed. More information can be found in the summary list.

The summary performs all calculations and its value is a list of the following:
\begin{itemize}
\item call
\item the variance estimate(s)
\item a matrix with the predictor in the first column, the estimated coefficient in the second column, the standard error in the third column, the z value in the fourth column, and the two-sided pvalue  in the fifth column. All inference is asymptotic, so we use the standard normal distribution to calculate the p-values.
\item the evaluated Monte Carlo log likelihood along with its first and second derivative
\item maybe other things that I haven't thought of
\end{itemize}

@@

Third and lastly, we want to show that 
\begin{align}
\sum_{k=1}^m \nabla \log \tilde{f} (u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}
\end{align}
 has finite variance. That is, we want to show the existence of
\begin{align}
\int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du.
\end{align}
We see that
\begin{align}
\int& \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du\\
&\leq  \int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)} du\\
&= \dfrac{1}{p_1} \int N(u|0,D) \left[\nabla \log \tilde{f}u) \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)+p_2 N(u|u^*,D^*)+p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})}   \right]^2 du\\
&\leq\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)}   \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{ \dfrac{\partial}{\partial \nu} N(u|0,D)}{ N(u|0,D)}   \right]^2 du\\
&= \dfrac{1}{p_1} \int \dfrac{\left[ \dfrac{\partial}{\partial \nu} N(u|0,D)  \right]^2}{N(u|0,D)} du \\
\end{align}
Now we use
\begin{align}
\dfrac{\partial}{\partial \nu} \log N(u|0,D) = \dfrac{\dfrac{\partial}{\partial \nu}  N(u|0,D)}{N(u|0,D)}
\end{align}
to see
\begin{align}
 \dfrac{1}{p_1} \int \dfrac{\left[ \dfrac{\partial}{\partial \nu} N(u|0,D)  \right]^2}{N(u|0,D)} du &=\dfrac{1}{p_1} \int  N(u|0,D) \left[ \dfrac{\partial}{\partial \nu} \log N(u|0,D)  \right]^2 du \\
&= \dfrac{1}{p_1} \int  N(u|0,D) \left[ \dfrac{\partial}{\partial \nu} \log N(u|0,D)  \right]^2 du 
\end{align}
which exists and is finite.
